Mallows, Black-box combinatorial optimization, limited budget, ???

### 2020-10-22 paper/main.tex
paper

\newcommand{\Prob}{\ensuremath{p}}
\newcommand{\minit}{\ensuremath{m_\text{ini}}\xspace}
\newcommand{\FEmax}{\ensuremath{\textit{FE}_{\max}}}

@@

\subsection{Black-box Combinatorial Optimization under a limited budget}

Permutations are defined as bijections of the set $[n]$ \manuel{integer}{of integers} onto itself. The set of all permutations of $n$ items is denoted as $S_n$ and has cardinality $n!$. We denote permutations with Greek letters with exception of the identity permutation denoted as $e=\manuel{}{(}1, 2, 3, \dotsc,n\manuel{}{)}$. We denote the composition of $\sigma$ and $\pi$ as $\sigma\pi$ and the inverse of $\sigma$ as $\sigma^{-1}$, for which the relation $\sigma\sigma^{-1}=e$ always holds.

\subsection{CEGO}

@@

\subsection{Distributions over permutations}
Distributions over permutations~\cite{critchlow91} are functions that assign a probability value  $\Prob(\sigma)\in[0,1]$ to each permutation $\sigma \in S_n$. One of the most popular distributions is the Mallows Model (MM), which is considered as an analogous to the Gaussian distribution for permutations. The MM defines the probability of each permutation $\sigma$ as follows:
%
\begin{equation}
\Prob(\sigma)=\frac{\exp(-\theta d(\sigma, \sigma_0))}{\psi}
\end{equation}
%
with two parameters, $\theta$ and $\sigma_0$, \manuel{Permutation $\sigma_0$ a reference}{where $\sigma_0$ is the}  permutation that has the largest probability value, i.e., the mode of the distribution. The probability of every permutation $\sigma\in S_n$ decays exponentially as its distance $d(\sigma,\sigma_0)$ increases, and the dispersion parameter $\theta$  controls this decay. The distance $d(\sigma,\sigma_0)$ is the Kendall's $\tau$ distance. The normalization constant $\psi$ can be easily computed for the Kendall's $\tau$ distance as well as for the Hamming, Cayley and Ulam distances~\cite{IruCalLoz2016permallows}.

The learning process\MANUEL{what learning process?} is divide in two stages: first, we estimate the central permutation of the distribution, $\hat{\sigma}_0$ and, second, \manuel{compute}{estimate} the dispersion parameter, $\hat\theta$~\cite{IruCalLoz2016permallows}.

The exact maximum likelihood estimator (MLE) for $\sigma_0$\MANUEL{for $\sigma_0$ o for $\hat{\sigma}_0$?} is given by the well-known Kemeny ranking~\cite{DwoKumNao2001rank}. Unfortunately,  obtaining such ranking is NP-hard~\cite{DwoKumNao2001rank}. An alternative to the Kemeny ranking is the Borda ranking~\cite{AliMei2011kemeny}, which has several advantages: (i) it requires polynomial computational time; (ii) it  guaranties  high quality parameters as an estimator for the MM~\cite{CarProSha2013votes}; and (iii) in a general optimization perspective, Borda error is bounded~\cite{CopFleRur2010ordering}.

The Borda ranking of \manuel{sample $S$}{a sample of permutations $S \subset S_n$} is computed as follows. First, for each item $i$ compute its Borda score $B(i) =  \sum_{t\in S}  \sigma_t(i)$. Then, assign rank 1 to the item with the smallest Borda score, rank 2 to the item with the second smallest score and so on, i.e., order the items by Borda score increasingly.\MANUEL{sorry, I don't follow here, what is an item? What does $\sigma_t(i)$ means?}

\subsection{Unbalanced Borda}\label{sec:uborda}
We introduce now the core for our probabilistic algorithm, the uBorda algorithm. Following the convention of the community, the ranking returned by the uBorda algorithm will be denoted uBorda ranking. The uBorda algorithm is a recent generalization of Borda~\cite{???}.\MANUEL{This paper is missing} In uBorda, the score of item $i$ is defined as follows:
%
\begin{equation}
\begin{split}
B(i) =  \sum_{t\in S}  w(\sigma_t) \sigma_t(i),
\end{split}
\label{eq:uborda_score}
\end{equation}
%
and, the same as Borda, orders the items increasingly by their score.\MANUEL{what is $w()$?}

Intuitively, the uBorda ranking of sample $S$\MANUEL{What is $S$?} is equivalent to the Borda ranking of $S'$, being $S'$ an extension of $S$ where each each $\sigma\in S$ has been replicated proportionally to its weight $w(\sigma)$.\MANUEL{so $w$ is a weight function from $w\colon S_n \to \mathbb{R}$?} In this paper, we would like to replicate each permutation $\sigma$ a number of times that is proportional to $f(\sigma)$.\MANUEL{We have not defined $f$ but I guess it is fintess, however, we minimize $f$, so how you handle that? I guess we need to define the black-box optimization problem earlier than here} Therefore, we will use $w(\sigma)=\rho^{f(\sigma)}$ for a $\rho\in(0,1)$.
%
By setting $\rho=1$ the original Borda is recovered. However, for $\rho<1$ uBorda ranking will be closer to the rankings in $S$ with smaller value of $f(\sigma)$.\MANUEL{by ``with smaller value'' do you mean ranking for increasing values of $f(\sigma)$?} Interestingly, Borda is computed in polynomial time. Clearly, the choice of the weight\MANUEL{by weight you mean $\rho$?} depends on the relevant property of a ranking in a particular domain. For example, it has been shown to have nice theoretical properties for the online estimation of the parameters of a sample with concept drift.\MANUEL{where? Do we miss a citation here?}

@@

\subsection{Sampling step}
Sampling can be done efficiently~\cite{IruCalLoz2016permallows} given the parameters that describe the model, i.e., $\sigma_0$ and $\theta$. Given these two parameters, sampling a MM has been studied in \cite{Collas}\cite{IruCalLoz2016permallows}\cite{Regenwetter2004} where the computational complexities are $O(n\log n)$ and $O(n^2)$.

@@

 \caption{Mean fitness (and standard deviation) of the best solution found at the end of each run, over 10 independent runs for each problem instance. The 95\% confidence interval corresponds to the two independent samples t-test on the mean of the difference between the fitness of UMM minus the one of CEGO. Mean runtime is measured in hours. If a run went over 72 hours, of  CEGO was not able to reach 400 function evaluations (FEs) in 72 hours, it was terminated earlier. This was the case for all LOP instances shown.\label{tab:results}}

>>>>

@@

Permutations are defined as bijections of the set $[n]$ integer onto itself. The set of all permutations of $n$ items is denoted as $S_n$ and has cardinality $n!$. We denote permutations with Greek letters with exception of the identity permutation denoted as $e=1, 2, 3, \ldots,n$. We denote the composition of $\sigma$ and $\pi$ as $\sigma\pi$ and the inverse of $\sigma$ as $\sigma^{-1}$, for which the relation $\sigma\sigma_{-1}=e$ always holds.

Distributions over permutations~\cite{critchlow91} are functions that assign a probability value to each of the permutations in $S_n$, $p(\sigma)\in[0,1]$. One of the most popular distributions is the Mallows Model (MM), which is considered as an analogous to the Gaussian distribution for permutations. The MM defines the probability of each permutation $\sigma$ as follows:

\begin{equation}
p(\sigma)=\frac{\exp(-\theta d(\sigma, \sigma_0))}{\psi}
\end{equation}
with two parameters, $\theta$ and $\sigma_0$, Permutation $\sigma_0$ a reference permutation that has the largest probability value, i.e., the mode of the distribution. The probability of every permutation $\sigma\in S_n$ decays exponentially as its distance $d(\sigma,\sigma_0)$ increases, and $\theta$, the dispersion parameter controls this decay. The distance $d(\sigma,\sigma_0)$ is the Kendall's-$\tau$ distance. The normalization constant $\psi$ can be easily computed for the Kendall's-$\tau$ distance as well as for the Hamming, Cayley and Ulam distance~\cite{Irurozki2016b}.

The learning process is divide in two stages: first, we estimate the central permutation of the distribution, $\hat\sigma_0$ and, second, compute the dispersion parameter, $\hat\theta$~\cite{Irurozki2016b}.

The exact maximum likelihood estimator (MLE) for $\sigma_0$ is given by the well-known Kemeny ranking~\cite{Dwork:2001:RAM:371920.372165}. Unfortunately,  obteinig such ranking is NP-hard~\cite{Dwork:2001:RAM:371920.372165}. An alternative to the Kemeny ranking is the Borda ranking~\cite{Ali2011}, which has several advantages. First, Borda requires polynomial computational time. Second, Borda as an estimator for the MM is guarantied to obtain high quality parameters~\cite{Caragiannis2013}. Finally, in a general optimization perspective, Borda error is bounded~\cite{Coppersmith:2010}.

The Borda ranking of sample $S$ is computed as follows: first, for each item $i$ compute its Borda score $B(i) =  \sum_{t\in S}  \sigma_t(i)$. Then, assign rank 1 to the item with the smallest Borda score, rank 2 to the item with the second smallest score and so on, i.e., order the items by Borda score increasingly.

\subsection{Unbalanced Borda}\label{sec:uborda}
In these lines we introduce the core for our probabilisitc algorithm, the uBorda algorithm. Following the convention of the community, the ranking returned by the uBorda algorithm will be denoted uBorda ranking. The uBorda algorithm is a recent generalization of Borda~\cite{}. In uBorda, the score of item $i$ is defined as follows:

\begin{equation}
\begin{split}
B(i) =  \sum_{t\in S}  w(\sigma_t) \sigma_t(i),
\end{split}
\label{eq:uborda_score}
\end{equation}
and, the same as Borda, orders the items increasingly by their score.

Intuitively, the uBorda ranking of sample $S$ is equivalent to the Borda ranking of $S'$, being $S'$ an extension of $S$ where each each $\sigma\in S$ has been replicated proportionally to its weight $w(\sigma)$. In this paper, we would like to replicate each permutation $\sigma$ a number of times that is proportional to $f(\sigma)$. Therefore, we will use $w(\sigma)=\rho^{f(\sigma)}$ for a $\rho\in(0,1)$.

By setting $\rho=1$ the original Borda is recovered. However, for $\rho<1$ uBorda ranking will be closer to the rankings in $S$ with smaller value of $f(\sigma)$. Interestingly, Borda is computed in polynomial time. Clearly, the choice of the weight depends on the relevant property of a ranking in a particular domain. For example, it has been shown to have nice theoretical properties for the online estimation of the parameters of a sample with concept drift.

\subsection{Previous method for black box optimization in combinatorial spaces: CEGO}

\manuel{CEGO: Manuel.}{}

@@

\newcommand{\minit}{\ensuremath{m_\text{ini}}\xspace}

\newcommand{\FEmax}{\ensuremath{\textit{FE}_{\max}}}

@@

\subsection{Sampling step}
Sampling can be done efficiently~\cite{Irurozki2016b} given the parameters that describe the model, i.e., $\sigma_0$ and $\theta$. Given these two parameters, sampling a MM has been studied in \cite{Collas}\cite{Irurozki2016b}\cite{Regenwetter2004} where the computational complexities are $O(n\log n)$ and $O(n^2)$.

@@

 \caption{Mean fitness (and standard deviation) of the best solution found at the end of each run, over 10 independent runs for each problem instance. The 95\% confidence interval corresponds to the two independent samples t-test on the mean of the difference between the fitness of UMM minus the one of CEGO. Mean runtime is measured in seconds. If a run of  CEGO was not able to reach 400 function evaluations (FEs) in 3 days, it was terminated earlier. This was the case for all LOP instances shown.\label{tab:results}}
